---
title: "Data Objects and Trip Records"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Objects and Trip Records}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(fsmr)
```

`fsmr` represents aggregate OD demand, zone attributes, and observed
origin-destination survey trips for a trip-based four-step workflow.

## Origin-Destination Demand

Use `fsm_od()` to standardize an OD table. Origin-destination pairs must be
unique. Demand can be zero or missing, but observed demand must be finite and
non-negative.

```{r od}
raw_demand <- data.frame(
  zone_o = c(1, 1, 2, 3),
  zone_d = c(2, 3, 1, 2),
  trips = c(100, 50, 200, 30)
)

od <- fsm_od(raw_demand, "zone_o", "zone_d", "trips")
od
summary(od)
plot(od)
```

`fsm_toy_od` is a ready-to-use OD example. It contains positive, zero, and
missing demand observations.

```{r toy-od}
fsm_toy_od
plot(fsm_toy_od)
```

## Zone Attributes

`fsm_zone()` stores a unique zone identifier and any attributes needed by a
model. Attribute names and types are user-defined.

```{r zones}
zone_data <- data.frame(
  id = c(1, 2, 3),
  residents = c(1200, 2500, 1800),
  employment = c(400, 1300, 900),
  school_places = c(150, 220, 80)
)

zones <- fsm_zone(zone_data, "id", od = od)
summary(zones)
plot(zones)
```

Supplying `od` validates that its origin and destination identifiers are known
zones. It does not require every zone to have observed trips.

## Origin-Destination Survey Trips

`fsm_trip()` stores one observed origin-destination survey trip per row.
Traveler and trip attributes retain their original names. An optional
`expansion_factor` represents the number of trips represented by a record.

```{r trips}
trip_data <- data.frame(
  id = 1:3,
  factor = c(1.2, 0.8, 1.5),
  income = c(1800, 2400, 0),
  employed = c(TRUE, TRUE, FALSE),
  student = c(FALSE, TRUE, FALSE),
  vehicles = c(TRUE, TRUE, FALSE),
  origin = c(1, 2, 1),
  destination = c(2, 1, 3),
  mode = c("car", "ridehailing", "walk+transit"),
  purpose_pair = c("home-work", "work-home", "home-health"),
  time_of_day = c("07:30:00", "17:15:00", "12:00:00")
)

trips <- fsm_trip(
  trip_data,
  trip_id = "id",
  expansion_factor = "factor"
)
trips
summary(trips)
```

`fsm_toy_trip` uses unit expansion factors and reproduces the positive-demand
records in `fsm_toy_od`. Its individuals make continuous trip chains, averaging
about 2.4 trips each.

```{r toy-trips, fig.width = 12, fig.height = 14}
fsm_toy_trip
plot(fsm_toy_trip)
```

## Aggregating Trip Records

Use the trip helpers to derive aggregate objects. For an `fsm_trip`,
`expansion_factor` is the default trip weight when it is present.

```{r trip-helpers}
trip_od <- fsm_od_from_trips(fsm_toy_trip)
transit_od <- fsm_od_from_trips(fsm_toy_trip, modes = "transit")
trip_zones <- fsm_zone_from_trips(fsm_toy_trip, include = 7L)
trip_totals <- fsm_od_totals_from_trips(
  fsm_toy_trip,
  zones = trip_zones
)

trip_od
transit_od
trip_totals
```

For a multimodal label such as `walk+transit`, filtering with `modes =
"transit"` matches the transit component.
