---
title: "Data Objects and Trip Records"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Objects and Trip Records}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(fsmr)
```

`fsmr` represents the aggregate and disaggregate inputs used in a trip-based
four-step workflow. It currently treats each origin-destination trip as an
independent choice unit.

## Origin-Destination Demand

Use `fsm_od()` to standardize an OD table. Origin-destination pairs must be
unique. Demand can be zero or missing, but observed demand must be finite and
non-negative.

```{r od}
raw_demand <- data.frame(
  zone_o = c(1, 1, 2, 3),
  zone_d = c(2, 3, 1, 2),
  trips = c(100, 50, 200, 30)
)

od <- fsm_od(raw_demand, "zone_o", "zone_d", "trips")
od
summary(od)
plot(od)
```

`fsm_toy_od` is a ready-to-use OD example. It contains positive, zero, and
missing demand observations.

```{r toy-od}
fsm_toy_od
plot(fsm_toy_od)
```

## Zone Attributes

`fsm_zone()` stores a unique zone identifier and any attributes needed by a
model. Attribute names and types are user-defined.

```{r zones}
zone_data <- data.frame(
  id = c(1, 2, 3),
  residents = c(1200, 2500, 1800),
  employment = c(400, 1300, 900),
  school_places = c(150, 220, 80)
)

zones <- fsm_zone(zone_data, "id", od = od)
summary(zones)
plot(zones)
```

Supplying `od` validates that its origin and destination identifiers are known
zones. It does not require every zone to have observed trips.

## Disaggregate Population and Trip Data

`fsm_population()` stores one disaggregate record per row. A record can be an
individual or a trip carrying individual attributes. `population_count`
represents multiple records only when every stored attribute is identical.

```{r population}
population_data <- data.frame(
  id = 1:3,
  people = c(2, 1, 3),
  income = c(1800, 2400, 0),
  employed = c(TRUE, TRUE, FALSE),
  student = c(FALSE, TRUE, FALSE),
  vehicles = c(TRUE, TRUE, FALSE),
  origin = c(1, 2, 1),
  destination = c(2, 1, 3),
  mode = c("car", "ridehailing", "walk+transit"),
  purpose_pair = c("home-work", "work-home", "home-health"),
  time_of_day = c("07:30:00", "17:15:00", "12:00:00")
)

population <- fsm_population(
  population_data,
  population_id = "id",
  population_count = "people"
)
population
```

`fsm_toy_population` uses trips as its observational unit. It keeps a
list-column of individual identifiers when otherwise identical trip records
are grouped. Its trip attributes include endpoints, mode, an ordered purpose
pair, and a 24-hour clock time.

```{r toy-population, fig.width = 12, fig.height = 14}
fsm_toy_population
summary(fsm_toy_population)
plot(fsm_toy_population)
```

## Aggregating Trip Records

Use the trip helpers to derive aggregate objects. For an `fsm_population`,
`population_count` is used as the default trip weight.

```{r trip-helpers}
trip_od <- fsm_od_from_trips(fsm_toy_population)
transit_od <- fsm_od_from_trips(fsm_toy_population, modes = "transit")
trip_zones <- fsm_zone_from_trips(fsm_toy_population, include = 7L)
trip_totals <- fsm_od_totals_from_trips(
  fsm_toy_population,
  zones = trip_zones
)

trip_od
transit_od
trip_totals
```

For a multimodal label such as `walk+transit`, filtering with `modes =
"transit"` matches the transit component.
