---
title: "Matrix Re-Aggregation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Matrix Re-Aggregation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(fsmr)
```

`fsm_reaggregate_matrix()` sums an origin-destination matrix into a different
zoning system. Supply a correspondence table with one row per source zone and
its target zone.

## A Simple Correspondence

```{r source-matrix}
matrix_od <- matrix(1:16, nrow = 4)
rownames(matrix_od) <- colnames(matrix_od) <- c("a", "b", "c", "d")
matrix_od

correspondence <- data.frame(
  old_zone = c("a", "b", "c", "d"),
  new_zone = c("north", "north", "south", "south")
)

fsm_reaggregate_matrix(matrix_od, correspondence)
```

The function applies the correspondence to both the origin and destination
dimensions. In this example, the four-zone matrix becomes a two-zone matrix.

## Separate Origin and Destination Correspondence

Origin and destination systems can use different correspondence columns.

```{r separate-correspondence}
row_map <- data.frame(
  source_zone = c("a", "b", "c", "d"),
  origin_district = c("west", "west", "east", "east")
)
col_map <- data.frame(
  source_zone = c("a", "b", "c", "d"),
  destination_district = c("north", "south", "north", "south")
)

fsm_reaggregate_matrix(
  matrix_od,
  row_map = row_map,
  col_map = col_map,
  from = "source_zone",
  to = "origin_district",
  col_from = "source_zone",
  col_to = "destination_district"
)
```

Every source row and column label must have a correspondence. The resulting
matrix preserves the total demand in the input matrix.

```{r total-check}
result <- fsm_reaggregate_matrix(matrix_od,
  row_map = row_map,
  col_map = col_map,
  from = "source_zone",
  to = "origin_district",
  col_from = "source_zone",
  col_to = "destination_district"
)

sum(matrix_od)
sum(result)
```
