---
title: "Trip Generation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Trip Generation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(fsmr)
```

Trip generation estimates productions and attractions. In an OD-based model,
observed productions are origin totals and observed attractions are destination
totals.

```{r observed-totals}
zones <- fsm_toy_zone
od <- fsm_toy_od
totals <- fsm_od_totals(od, zones)

totals
summary(totals)
plot(totals)
```

## Regression

`fsm_generation()` fits production and attraction regressions independently.
Supply one formula for a one-sided model or both formulas for a two-sided
model. Predictions are rounded to whole trips.

```{r regression}
fit <- fsm_generation(
  zones,
  totals,
  method = "regression",
  production_formula = production ~ 0 + population,
  attraction_formula = attraction ~ 0 + jobs
)

summary(fit)
predict(fit)
plot(fit)
```

Coefficient confidence intervals are available for regression fits. The
result is a named list because production and attraction are separate models.

```{r confidence-intervals}
intervals <- confint(fit)
intervals$production
intervals$attraction
```

Use `newdata` to apply the fitted model elsewhere. Missing predictors produce
`NA` only for the affected outcome. The `negative` argument controls negative
predictions.

```{r prediction-controls}
predict(fit, zones, negative = "zero")
predict(fit, zones, negative = "keep")

production_only <- fsm_generation(
  zones,
  totals[, c("zone_id", "production"), with = FALSE],
  method = "regression",
  production_formula = production ~ 0 + population
)
predict(production_only, zones)
```

## Cross-Classification

Cross-classification estimates separate production and attraction rates for
each combination of user-selected attributes. Exposure columns transform those
rates into trip totals.

```{r classification}
rate_fit <- fsm_generation(
  zones,
  totals,
  method = "cross_classification",
  cross_classification = c("area_type", "income_group"),
  production_exposure = "households",
  attraction_exposure = "jobs"
)

summary(rate_fit)
predict(rate_fit, zones)
```

The fitted cells are available from the summary object.

```{r classification-rates}
summary(rate_fit)$cell_rates
```

## Growth Factors

Growth factors scale observed totals. A scalar model can be stored as a
reusable rule without data or observed totals.

```{r growth}
growth_fit <- fsm_generation(
  zones,
  totals,
  method = "growth_factor",
  growth_factor = 1.10
)
predict(growth_fit)

reusable_growth <- fsm_generation(
  method = "growth_factor",
  growth_factor = 1.10
)
predict(reusable_growth, totals)
```

Zone-specific growth factors can change systemwide production and attraction
by different amounts because their starting zonal distributions differ.

## Custom and Externally Calibrated Models

Use `custom_fit` and `custom_predict` for a model fitted inside the workflow.

```{r custom}
custom_fit <- fsm_generation(
  zones,
  totals,
  method = "custom",
  custom_fit = function(data, totals, ...) {
    list(
      production = mean(totals$production),
      attraction = mean(totals$attraction)
    )
  },
  custom_predict = function(object, newdata, ...) {
    data.frame(
      zone_id = newdata$zone_id,
      production = object$production,
      attraction = object$attraction
    )
  }
)
predict(custom_fit, zones)
```

For parameters calibrated outside `fsmr`, supply `custom_object` directly.
No observed totals or dummy fitting function are needed.

```{r external}
parameters <- list(
  production = c(intercept = 100, population = 0.04),
  attraction = c(intercept = 50, jobs = 0.06)
)

external_fit <- fsm_generation(
  method = "custom",
  custom_object = parameters,
  custom_predict = function(object, newdata, ...) {
    data.frame(
      zone_id = newdata$zone_id,
      production = object$production[["intercept"]] +
        object$production[["population"]] * newdata$population,
      attraction = object$attraction[["intercept"]] +
        object$attraction[["jobs"]] * newdata$jobs
    )
  }
)
predict(external_fit, zones)
```

## Trip-Level Records

Trip-level generation uses `population_id` as the record identifier. The toy
population includes individual and trip attributes, and `population_count`
acts as the record weight.

```{r trip-level}
population_totals <- data.frame(
  population_id = fsm_toy_population$population_id,
  production = fsm_toy_population$population_count *
    (1 + fsm_toy_population$employed + fsm_toy_population$vehicles),
  attraction = fsm_toy_population$population_count *
    (2 + 2 * fsm_toy_population$employed + fsm_toy_population$vehicles)
)

population_fit <- fsm_generation(
  fsm_toy_population,
  population_totals,
  method = "regression",
  production_formula = production ~ 0 + population_count +
    population_count:employed + population_count:vehicles,
  attraction_formula = attraction ~ 0 + population_count +
    population_count:employed + population_count:vehicles
)

predict(population_fit)
```
